---
title: "Analysis"
output:  
  html_document:
    number_sections: true
    toc: true  
    collapsed: false
    toc_float: true
    smooth_scroll: false
    toc_depth: 3
---




```{r, message = FALSE, echo = FALSE, warning=FALSE}
# packages
library(reshape2)
library(ggplot2)
library(plyr)
library(ez)
library(gridExtra)
library(cowplot)
library(lme4)
library(effects)
library(multcomp)
library(psych)
library(gmodels)


library(showtext)
library(readr)


font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()

# function
analyzeLogModel<-function(logModel){
  dev<-logModel$deviance
  nullDev<-logModel$null.deviance
  df<-logModel$df.residual
  nullDf<-logModel$df.null
  modelN<-length(logModel$fitted.values)
  modelChi<-nullDev-dev
  chiDf<-nullDf-df
  chiProb<-1-pchisq(modelChi,chiDf)
  R.l<-1-dev/nullDev
  R.cs<-1-exp(-(nullDev-dev)/modelN)
  R.n<-R.cs/(1-(exp(-(nullDev/modelN))))
  OR<-exp(logModel$coefficients)
  cat("Model Chi^2             ",round(modelChi,3),"\n")
  cat("df                      ",chiDf,"\n")
  cat("p                       ",round(chiProb,5),"\n")
  cat("Hosmer and Lemeshow R^2 ",round(R.l,3),"\n")
  cat("Cox and Snell R^2       ",round(R.cs,3),"\n")
  cat("Nagelkerke R^2          ",round(R.n,3),"\n")
  cat("Odds Ratios             ",round(OR,3),"\n")
}

```


# Data preparation

```{r}
tdata <- read.delim("03_tdata.txt",header=FALSE)
sdata <- read.delim("03_sdata.txt",header=FALSE)
cond.lookup <- read.delim("03_cond_lookup.txt",header=TRUE)


# reshape data -----------------------------------------------------------

names(tdata)<-c("sID","order","learn","colorHeavy","side","variable","value")
names(sdata)<-c("sID","sex","age","duration","code","dispcode","size_1","size_2","weight_1","weight_2","attCheck","alltag","email")
tdata <- tdata[with(tdata, order(sID)),]
sdata <- sdata[with(sdata, order(sID)),]
tdata$sID <- as.factor(tdata$sID)
tdata$value <- tdata$value-1
sdata <- subset(sdata,subset=sID!=569) # subject was not assigned to any condition due to technical error

### generate subject data in wide format

tdata.wide <- dcast(tdata,sID+order+learn+colorHeavy+side~variable)
data.wide <- join(tdata.wide,sdata,by="sID")

# evaluate whether observations were correct

for (i in 1:dim(data.wide)[1]){
  if((data.wide$colorHeavy[i]=="orange"&data.wide$obs_orange[i]==0&data.wide$obs_violet[i]==1)|
     (data.wide$colorHeavy[i]=="violet"&data.wide$obs_orange[i]==1&data.wide$obs_violet[i]==0)) data.wide$obs_corr[i] <- 1
  else data.wide$obs_corr[i] <- 0
}

# evaluate whether manipulation checks were correct

data.wide$size_corr <- abs(data.wide$size_1-2)

for (i in 1:dim(data.wide)[1]){
  if ((data.wide$weight_1[i]==2&data.wide$colorHeavy[i]=="orange"&data.wide$weight_2[i]==1)|
      (data.wide$weight_1[i]==2&data.wide$colorHeavy[i]=="violet"&data.wide$weight_2[i]==2)) data.wide$weight_corr[i] <- 1
  else data.wide$weight_corr[i] <- 0
}

for (i in 1:dim(data.wide)[1]){
  if (data.wide$weight_corr[i]==1&data.wide$size_corr[i]==1) data.wide$mc_corr[i] <- 1
  else data.wide$mc_corr[i] <- 0
}

# evaluate whether subjects gave correct responses on observation questions and attention check

for (i in 1:dim(data.wide)[1]){
  if (data.wide$obs_corr[i]==1&data.wide$attCheck[i]==2) data.wide$valid[i] <- 1
  else data.wide$valid[i] <- 0
}

### generate choice responses in long format

# evaluate whether intended prediction was made using lookup table

tdata.choice <- subset(tdata, subset = substr(variable,10,15)=="choice")
tdata.choice$ind <- paste(tdata.choice$colorHeavy,tdata.choice$side,substr(tdata.choice$variable,1,8),sep="_")
tdata.choice <- join(tdata.choice,cond.lookup,by="ind")

for (i in 1:dim(tdata.choice)[1]){
  if(tdata.choice$value[i]==tdata.choice$intAns[i]) tdata.choice$choice_corr[i] <- 1
  else tdata.choice$choice_corr[i] <- 0
}

tdata.choice$variable <- factor(tdata.choice$variable)

# code test machines

for (i in 1:dim(tdata.choice)[1]){
  if(tdata.choice$learn[i]==substr(tdata.choice$variable[i],1,1)) tdata.choice$test_machine[i] <- "identical"
  else if (substr(tdata.choice$variable[i],1,1)=="C") tdata.choice$test_machine[i] <- "transfer_2"
  else tdata.choice$test_machine[i] <- "transfer_1"
}

# code test object

for (i in 1:dim(tdata.choice)[1]){
  if(substr(tdata.choice$ind[i],1,6)==substr(tdata.choice$ind[i],15,20)) tdata.choice$test_object[i] <- "heavy"
  else tdata.choice$test_object[i] <- "light"
}

tdata.choice <- join(tdata.choice,subset(data.wide,select=c(sID,obs_corr,mc_corr,valid)),by="sID")

### generate certainty responses in long format

tdata.certainty <- subset(tdata, subset = substr(variable,10,16)=="certain")
for (i in 1:dim(tdata.certainty)[1]){
  if(tdata.certainty$learn[i]==substr(tdata.certainty$variable[i],1,1)) tdata.certainty$test_machine[i] <- "identical"
  else if (substr(tdata.certainty$variable[i],1,1)=="C") tdata.certainty$test_machine[i] <- "transfer_2"
  else tdata.certainty$test_machine[i] <- "transfer_1"
}
tdata.certainty$test_object <- tdata.choice$test_object
tdata.certainty <- join(tdata.certainty,subset(data.wide,select=c(sID,obs_corr,mc_corr,valid)),by="sID")
tdata.certainty$value <- tdata.certainty$value+1

### combine all tables into one wide format dataframe

data.complete <- join(data.wide,dcast(subset(tdata.choice,select = c(sID,test_machine,test_object,choice_corr)),sID~test_machine+test_object),by="sID")
data.complete <- join(data.complete,dcast(subset(tdata.certainty,select = c(sID,test_machine,test_object,value)),sID~test_machine+test_object),by="sID")
names(data.complete)[37:48] <- c("choice_identical_heavy","choice_identical_light","choice_transfer1_heavy","choice_transfer1_light","choice_transfer2_heavy","choice_transfer2_light",
                                 "certainty_identical_heavy","certainty_identical_light","certainty_transfer1_heavy","certainty_transfer1_light","certainty_transfer2_heavy","certainty_transfer2_light")

# write.table(data.complete, file="data_complete.txt", append=TRUE, sep="\t", col.names=TRUE, row.names=FALSE, quote=FALSE)
```

# Graphs

## Proportions

```{r}
# plot proportions --------------------------------------------------------

myTheme <- theme(plot.title = element_text(face="bold", size = 24),
        axis.title.x = element_text(size = 24),
        axis.title.y = element_text(size = 24),
        axis.text.x = element_text(size = 24, angle = 0), 
        axis.text.y = element_text(size = 24, angle = 0),
        legend.text = element_text(size = 24),
        legend.title = element_text(face = "bold", size =24),
        strip.text.x = element_text(size = 24),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))

theme_set(theme_light(base_size = 10, base_family = "Poppins"))


# data <- tdata.choice
data <- subset(tdata.choice,subset=valid==1)
# data <- subset(tdata.choice,subset=obs_corr==1)

# proportions by test machine and test object

df.prop <- ddply(data[,c(10:12)],.(test_machine,test_object), summarize,
                 prop=as.numeric(prop.table(table(choice_corr))),
                 opt=as.factor(names(table(choice_corr))))
df.prop <- df.prop[df.prop$opt==1,-4]
df.prop$n <- ddply(data[,c(10:12)], .(test_machine,test_object), function(df)length(df$choice_corr))[,3]
df.prop$k <- df.prop$n*df.prop$prop

bar.prop <- ggplot(df.prop,aes(test_machine,prop,fill=test_object))
bar.prop +
  coord_cartesian(ylim=c(0,1)) +
  scale_y_continuous(limits = c(0,1), breaks=seq(0, 1, .1)) +
  scale_x_discrete(labels=c("Identical", "Rearrangement", "Novel")) +
  geom_bar(stat="identity", position='dodge', colour = "black") + # stat="summary", fun.y = mean for average across factors
  geom_hline(data=NULL, aes(yintercept=.5), linetype = "dashed", size=1)+
  myTheme+
  #theme(plot.title = element_text(face="bold", size = 12),
   #     axis.title.x = element_text(face = "bold", size = 12),
    #    axis.title.y = element_text(face = "bold", size = 12),
     #   axis.text.x = element_text(size = 12), 
      #  axis.text.y = element_text(size = 12),
       # legend.text = element_text(size = 12),
        #legend.title = element_text(face = "bold", size = 12),
        #strip.text.x = element_text(size = 12),
        #strip.background = element_rect(colour="black", fill="light grey"),
        #legend.position=c(0.5,0.25),
        #         legend.title=element_blank(), 
        #panel.grid.major = element_blank(), 
        #panel.grid.minor = element_blank(), 
        #panel.background = element_blank(), 
        #axis.line = element_line(colour = "black"), 
        #axis.text = element_text(colour ="black"), 
        #axis.ticks = element_line(colour ="black")) + 
  labs(x = "Transfer Condition", y = "Proportion") +
  scale_fill_brewer(palette = "Pastel1", 
                    name="Test Object", labels=c("Heavy","Light"))

 ggsave("Figure_1.png",width=5,height=3)
```


## Certainty ratings 

```{r}
# plot certainty ratings --------------------------------------------------

data <- tdata.certainty

df.summary <- ddply(data[,c(7:9)],.(test_machine,test_object),function(df)c(mean(df$value),sd(df$value),length(df$value))) # ddply is from plyr
names(df.summary) <- c("machine","object","mean","sd","n")
df.summary$MOE.95 <- (df.summary$sd/sqrt(df.summary$n))*qnorm(.975,0,1)
df.summary$UB.95 <- df.summary$mean+df.summary$MOE.95
df.summary$LB.95 <- df.summary$mean-df.summary$MOE.95

bar.summary <- ggplot(df.summary,aes(machine,mean,fill=object))
bar.summary+ 
  coord_cartesian(ylim=c(1,5))+ 
  scale_y_continuous(limits=c(0,5),breaks=seq(1,5,1))+
  scale_x_discrete(labels=c("Identical", "Rearrangement", "Novel")) +
  geom_bar(stat="identity",position="dodge",colour="black")+
  geom_errorbar(mapping=aes(ymin=LB.95,ymax=UB.95),stat="identity",position=position_dodge(width=.9),width=.2)+
#   ggtitle("Title")+
  myTheme+
  #theme(plot.title=element_text(face="bold",size = 12),
    #    axis.title.x=element_text(face="bold",size=12),
     #   axis.title.y=element_text(face="bold",size=12),
      #  axis.text.x=element_text(size=12), 
       # axis.text.y=element_text(size=12),
        #legend.text=element_text(size=12),
        #legend.title=element_text(face="bold",size=12),
        #strip.text.x=element_text(size=12),
        #strip.background=element_rect(colour="black",fill="light grey"),
        #legend.position=c(0.5,0.25),
        #panel.grid.major=element_blank(), 
        #panel.grid.minor=element_blank(), 
        #panel.background=element_blank(), 
        #axis.line=element_line(colour="black"), 
        #axis.text=element_text(colour="black"), 
        #axis.ticks=element_line(colour="black")) + 
  labs(x="Transfer Condition",y="Certainty") +
  scale_fill_manual(values=c("#696969","#999999"),name="Test Object",labels=c("Heavy","Light"))

ggsave("Figure_2.png",width=5,height=3)

```

# Statistical analysis

## Descriptive stats
```{r}
# statistics --------------------------------------------------------------

data <- subset(tdata.choice,subset=valid==1,select=c("sID","test_machine","test_object","choice_corr"))

data$test_machine <- factor(data$test_machine)
data$test_object <- factor(data$test_object)

by(data$choice_corr,data[,c(3,2)],describe)
```

## Chi-squared tests against chance performance by group 
```{r}
### Chi-squared tests against chance performance by group

df.prop <- ddply(data[,c(2:4)],.(test_machine,test_object), summarize,
                 prop=as.numeric(prop.table(table(choice_corr))),
                 opt=as.factor(names(table(choice_corr))))
df.prop <- df.prop[df.prop$opt==1,-4]
df.prop$n <- ddply(data[,c(2:4)], .(test_machine,test_object), function(df)length(df$choice_corr))[,3]
df.prop$k <- df.prop$n*df.prop$prop

chisq.vs.5 <- list()
for (i in 1:dim(df.prop)[1]){
  chisq.vs.5[[i]]<-chisq.test(c(df.prop$k[i],df.prop$n[i]-df.prop$k[i]))
}
chisq.vs.5<-data.frame(do.call(rbind,chisq.vs.5))[,c(1:3)]
chisq.vs.5$statistic<-round(as.numeric(chisq.vs.5$statistic),3)
chisq.vs.5$parameter<-as.numeric(chisq.vs.5$parameter)
chisq.vs.5$p.value<-round(as.numeric(chisq.vs.5$p.value),3)
chisq.vs.5[,c(4:6)]<-df.prop[,c(1,2,4)]
chisq.vs.5<-chisq.vs.5[,c(4,5,2,1,3)]

chisq.vs.5

write.table(chisq.vs.5,file="chi_test.txt",append=TRUE,sep="\t",col.names=TRUE,row.names=FALSE,quote=FALSE)
```

## Generalized linear mixed model: Main analysis
```{r}
### Generalized linear mixed model: Main analysis

data <- subset(tdata.choice,subset=valid==1,select=c("sID","test_machine","test_object","choice_corr","order","learn","colorHeavy","side"))

data$test_object<-as.factor(data$test_object)
data$test_machine<-as.factor(data$test_machine)

model.0<-glmer(choice_corr~1+(1|sID),family=binomial(),data=data)
model.tm<-glmer(choice_corr~test_machine+(1|sID),family=binomial(),data=data)
model.additiv<-glmer(choice_corr~test_machine+test_object+(1|sID),family=binomial(),data=data)
model.full<-glmer(choice_corr~test_machine*test_object+(1|sID),family=binomial(),data=data)

anova(model.0,model.tm,model.additiv,model.full)
```

## Generalized linear mixed model: Balancing factors
```{r}
### Generalized linear mixed model: Balancing factors

model.bal<-glmer(choice_corr~order+learn+colorHeavy+side+(1|sID),family=binomial(),data=data)
summary(model.bal)
```

Summary of the analysis of the counterbalancing factors: 

With the analysis above, we check whether the different balancing procedures affected performance in the task. The generalized linear mixed model predicts performance by learning device, device order, color, and side (not allowing for interactions) across all trial types. The results table indicates that device order and color did not affect the probability of correct responding, β = .11, z = .34, p = .74 and β = .03, z = .08, p = .94, respectively. Learning with device B (see-saw) slightly reduced performance relative to learning with device A (chute), β = -.72, z = -2.14, p < .05. Apparently, the regularity information conveyed via device A, in which one of the objects removed an obstacle, allowed for more consensual capacity inferences than device B, in which one of the objects outbalanced a see-saw. Finally, the side factor also affected performance, β = -.71, z = -2.10, p < .05. The probability of correct responding was slightly reduced when the test devices were a mirror image of the learning devices, as opposed to when the devices were not mirrored. This might indicate that a few subjects simply matched their prediction to the concrete outcome observed during learning.


## ANOVA for certainty ratings
```{r}
### ANOVA for certainty ratings

anova.certainty <- ezANOVA(subset(tdata.certainty,subset=valid==1), dv = .(value), wid = .(sID), within = .(test_machine,test_object), type = 3, detailed = TRUE) # ezANOVA() is from ez
anova.certainty
```

